---
title: "Stacked marplot"
author: "Tim Johnson"
date: "4/5/2018"
output: html_document
---

OK so now we move forward to using phyloseq to investigate the taxonomic composition of the different communities.

```{r setup}
knitr::opts_knit$set(root.dir = '~/USDA_laptop//Desktop/BMD_turkeys/data/metabolome/')

#Load libraries
library(ggplot2)
library(vegan)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(tidyr)
```

```{r}
#Assign varibles for the paths of the data to import
sharedfile <- "bmd_only.tax.shared"
taxfile <- "../stability.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.abund.rdp.wang.tx.1.cons.taxonomy"
mapfile <- "../../BMD_design.txt"

#Import mothur data
mothur_data <- import_mothur(mothur_shared_file = sharedfile, mothur_constaxonomy_file = taxfile)
head(meta)
meta$day <- factor(meta$day)
```

We convert this dataframe into phyloseq format with a simple constructor `sample_data`. The only formatting required to merge the sample data into a phyloseq object is that the rownames must match the sample names in your shared and taxonomy files.

```{r}
metadata <- sample_data(meta)
rownames(metadata) <- metadata$name

#Merge this into our phyloseq object
BMD_data <- merge_phyloseq(mothur_data, metadata)
BMD_data
```
Lets check some things to make sure everything looks alright. First taxonomy level names

```{r}
colnames(tax_table(BMD_data))

colnames(tax_table(BMD_data)) <- c("Kingdom", "Phylum", "Class", 
  "Order", "Family", "Genus")
```

This is also a good time to check all your samples and taxa to make sure you dont need to filter out PCR negative controls or mitochondria or chloroplast OTUs out of your dataset. In my case this has already been done.

So now we will make some beautiful stacked barplots to get an idea of what is going on in these samples.

```{r}
# Set colors for plotting
my_colors <- c(
  '#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c',
  '#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#ffff99','#b15928', 
  "#CBD588", "#5F7FC7", "orange","#DA5724", "#508578", "#CD9BCD",
   "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861", "black"
)

#If you want different taxonomic level, find and replace the taxonomic level listed here
my_level <- "Family"
rm(bmd.summary)
bmd.summary <- BMD_data %>%
  tax_glom(taxrank = my_level) %>%                     # agglomerate at Family level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt()  %>%                                           # Melt to long format
  group_by(group, Family) %>%
  summarise(Abundance.average=mean(Abundance)) 

bmd.taxa.average <- bmd.summary %>% 
  group_by(Family) %>%
  summarise(Family.average=mean(Abundance.average))
  
# merging the phyla means with the metadata #
bmd.summary <- merge(bmd.summary, bmd.taxa.average)
bmd_meta <- merge(bmd.summary, unique(meta[,-1]) , by = 'group')

abund_filter <- 0.01
bmd_meta2 <- filter(bmd_meta, Family.average>abund_filter)
str(bmd_meta2)
bmd_meta2$location <- factor(bmd_meta2$location, levels = c("jc", "ic", "cc"))
unique(bmd_meta2$Family)

# Plot 
ggplot(bmd_meta2, aes(x = trt, y = Abundance.average, fill = Family)) + 
  facet_grid(location~day) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = my_colors) +
  # Remove x axis title
  #theme(axis.title.x = element_blank()) + 
  ylim(c(0,1)) +
  guides(fill = guide_legend(reverse = F, keywidth = .5, keyheight = .5, ncol = 1)) +
  theme(legend.text=element_text(size=8)) +
  #theme(legend.position="bottom") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ylab(paste0("Relative Abundance (Family > ", abund_filter * 100,"%) \n")) +
  ggtitle("Family Composition of turkey GI tract \n and impact of BMD treatment") 
ggsave("../../analysis/Family2BarPlot_AllSamples.png", height = 5)


```
```{r}

# Plot 
ggplot(bmd_meta2, aes(x = day, y = Abundance.average, fill = Family)) + 
  facet_grid(location~trt) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = my_colors) +
  # Remove x axis title
  #theme(axis.title.x = element_blank()) + 
  ylim(c(0,1)) +
  guides(fill = guide_legend(reverse = F, keywidth = .5, keyheight = .5, ncol = 1)) +
  theme(legend.text=element_text(size=8)) +
  #theme(legend.position="bottom") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ylab(paste0("Relative Abundance (Family > ", abund_filter * 100,"%) \n")) +
  ggtitle("Family Composition of turkey GI tract \n and impact of BMD treatment") 
ggsave("../../analysis/FamilyTimeBarPlot_AllSamples.png", height = 5)

```